#"Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true



## Loading and preprocessing the data
```{r}
setwd("/Users/hitanshupande/Documents/Reproducible Research/")
library(dplyr)
library(ggplot2)
data <- read.csv("activity.csv")
```


## Add a new column with the type of day, weekday or weekend 
```{r}
data$weekday <- as.POSIXct(data$date)
data$weekday <- weekdays(data$weekday)
```

## Remove NA values
```{r}
cleandata <- data[!is.na(data$steps),]
```

## calculate sum of all steps by date
```{r}

steps.date <- aggregate(steps ~ date, cleandata, sum)
```

##Plot a histogram of total steps taken
```{r}
hist(steps.date$steps)
```


## What is mean total number of steps taken per day?

##Mean of steps taken per day
```{r}
mean(steps.date$steps)
```

##Median of steps taken per day
```{r}
median(steps.date$steps)
```

## What is the average daily activity pattern?

##Calculate sum of steps per interval
```{r}
intervalAvg <- aggregate(steps ~ interval, cleandata, sum)
```

##Plot time series graph of steps per interval
```{r}
plot.ts(intervalAvg)
```

## Interval with maximum number of steps (835)
```{r}
intervalAvg[intervalAvg$steps==max(intervalAvg$steps),]
```

##Total number of NA values in dataset
```{r}
nrow(data[is.na(data$steps),])
```

##Strategy to fill NA values is to replace them with average steps for given time interval
##Create a dataframe with only NA values
```{r}
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]

steps_full <- data_full %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
## Plot histogram 
    ggplot(steps_full, aes(x = steps)) +
    geom_histogram(fill = "firebrick", binwidth = 1000) +
    labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")

mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)

mean_steps_full
median_steps_full
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_full <- mutate(data_full, weekday = ifelse(data_full$date == "Saturday" | data_full$date == "Sunday", "weekend", "weekday"))
data_full$weekday <- as.factor(data_full$weekday)
head(data_full)
tail(data_full)
data_full
interval_full <- data_full %>%
    group_by(interval, weekday) %>%
    summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weekday)) +
    geom_line()
print(s)
```
